%\documentclass{hotnets22}
% \documentclass[sigconf,10pt]{acmart}
\documentclass[letterpaper,twocolumn,10pt]{article}
\usepackage{usenix-2020-09}
\usepackage{amsmath}
\usepackage{amsthm}

\usepackage{xfrac}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{times}  
\usepackage{hyperref}
% \usepackage{subfig}
\usepackage{tikz}
\usetikzlibrary{math}
\usepackage{pgfplots}
\usetikzlibrary{pgfplots.groupplots}
\usepackage{pgfplotstable}
\usepackage[subtle]{savetrees}
\usepackage[title]{appendix}

% axis style, ticks, etc
\pgfplotsset{every axis/.append style={
                    label style={font=\footnotesize},
                    tick label style={font=\footnotesize}  
                    }}

\usepackage{caption}
\usepackage{subcaption}
\usepackage{multirow}
\usepackage{tabularx}
\usepackage{array}
\usepackage{xspace}
\usepackage{soul}
\usepackage[normalem]{ulem}

\newcolumntype{s}{>{\hsize=.3\hsize\linewidth=\hsize}X}
\newcolumntype{D}{>{\hsize=.3\hsize\linewidth=\hsize}X}

\newcommand{\wdImg}{\dimexpr \linewidth-2\tabcolsep} %width of the image

\hypersetup{pdfstartview=FitH,pdfpagelayout=SinglePage}

\setlength\paperheight {11in}
\setlength\paperwidth {8.5in}
\setlength{\textwidth}{7in}
\setlength{\textheight}{9.25in}
\setlength{\oddsidemargin}{-.25in}
\setlength{\evensidemargin}{-.25in}

\newcommand{\rg}[1]{\textcolor{blue}{(\textbf{GR:} #1)}}
\newcommand{\vk}[1]{\textcolor{green}{(\textbf{VA:} #1)}}

\newtheorem{theorem}{Theorem}
\newtheorem{definition}{Definition}
\newtheorem{lemma}{Lemma}
\newtheorem{observation}{Observation}

% we have a big fig
\renewcommand{\floatpagefraction}{.8}%
% \renewcommand{\topfraction}{.8}
% \renewcommand{\bottomfraction}{.8}
\newcommand{\nftables}{\texttt{nftables}\xspace} 

% TODO: remove for the CR!
% \pagestyle{plain}
% \settopmatter{printfolios=true}

\begin{document}

% \conferenceinfo{HotNets 2022} {}
% \CopyrightYear{2022}
% \crdata{X}
% \date{}

%%%%%%%%%%%% THIS IS WHERE WE PUT IN THE TITLE AND AUTHORS %%%%%%%%%%%%

\title{More Bang for the Buck: 
Superlinear Scaling with Distributed Self-adjusting Systems}

\author{Paper \#619, 12 + 7 pages}
% \author{Jonas Köppeler, Maciej Pacut, Tamás Lévai, Vamsi Addanki, Stefan Schmid, Gábor Rétvári}

\maketitle

\begin{abstract}
  Extracting maximum performance from a limited pool of parallel compute resources remains a central challenge.  In this paper, we show an optimization technique that allows certain distributed systems to attain faster-than-linear (superlinear) performance improvement with only a linear scaling of the worker pool. Our insight is that (1) dispatching jobs to parallel workers so that the locality of reference in the workers' input increases and (2) implementing the workers with a self-adjusting algorithm to take advantage of the higher locality can yield superlinear scaling in many practical applications. First, we demonstrate our technique in simulations: by scaling textbook self-adjusting algorithms, we obtain 100--3,300x speedup using only 48 CPU cores --- up to 70x beyond linear scaling. After that, we present operational case studies. As a major contribution, we re-engineer the default Linux packet classifier to attain 5--25x raw performance improvement as compared to the vanilla kernel. We demonstrate 800x speedup on synthetic traces and 220x speedup on real firewall traces with 32 CPU cores. We also provide a formal model and a set of design guidelines to help understand the applicability of our optimization strategy for particular distributed system workloads.
\end{abstract}


% We further apply our methodology to demonstrate a 2.3x faster than linear scaling on a Memcached/PostgreSQL storage system. 

% \tableofcontents

% a unique combination of \emph{locality-boosting load balancing} to spread load among workers implemented using \emph{self-adjusting data structures}.
  
\input{intro.tex}

\input{background.tex}

\input{architecture.tex}

\input{case-study-classifier.tex}

\input{discussion.tex}

\input{related.tex}

\input{conclusions.tex}

%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \input{ack.tex}

\bibliographystyle{abbrv} 
\begin{small}
\bibliography{mrf}
\end{small}

\begin{appendices}
\input{analysis.tex}

\input{case-study-dist-caching.tex}
\end{appendices}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
